---
title: "Assignment"
author: "Charles Cheruparambil"
date: "2023-03-23"
output: html_document
---
Load the libraries:
```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(rtracklayer)
  library(epiwraps)
  library(GenomicRanges)
})
ah <- AnnotationHub()
```

#Assignment

Find and download from ENCODE the peaks (i.e. bed-like format) for the following in mouse embryonic stem cells (mESC) :

-Transcription factor p300
-histone modifications H3K4me3, H3K4me1, H3K27ac, and H3K27me3
-> (when there are replicates, we recommend using the bed file denoted as “conservative IDR thresholded peaks”)

##Downloading the processed data
https://www.encodeproject.org/ -> Experiment Search -> apply Filter -> choose filetype 'bed narrow' of experiment summary -> import data:
###Transcription factor p300:
```{r}
download.file("https://www.encodeproject.org/files/ENCFF460EOA/@@download/ENCFF460EOA.bed.gz", dest="tf_p300.bed.gz")
tf_p300_peaks <- rtracklayer::import("tf_p300.bed.gz", format="narrowPeak")
tf_p300_peaks
```

-> GRanges object with 28341 ranges and 6 metadata columns -> total 28341 peaks.

### H3K4me3
```{r}
download.file("https://www.encodeproject.org/files/ENCFF671UNN/@@download/ENCFF671UNN.bed.gz", dest="hist_H3K4me3.bed.gz")
hist_H3K4me3_peaks <- rtracklayer::import("hist_H3K4me3.bed.gz", format="narrowPeak")
```
### H3K4me1
```{r}
download.file("https://www.encodeproject.org/files/ENCFF333IJH/@@download/ENCFF333IJH.bed.gz", dest="hist_H3K4me1.bed.gz")
hist_H3K4me1_peaks <- rtracklayer::import("hist_H3K4me1.bed.gz", format="narrowPeak")
```
### H3K27ac
```{r}
download.file("https://www.encodeproject.org/files/ENCFF857FNO/@@download/ENCFF857FNO.bed.gz", dest="hist_H3K27ac.bed.gz")
hist_H3K27ac_peaks <- rtracklayer::import("hist_H3K27ac.bed.gz", format="narrowPeak")
```

### H3K27me3
```{r}
download.file("https://www.encodeproject.org/files/ENCFF105NKG/@@download/ENCFF105NKG.bed.gz", dest="hist_H3K27me3.bed.gz")
hist_H3K27me3_peaks <- rtracklayer::import("hist_H3K27me3.bed.gz", format="narrowPeak")
```
##QUESTION: Of the p300 peaks, what proportion overlap each of the marks?

### H3K4me3

```{r}
peaksOverlapsH3K4me3 <- overlapsAny(tf_p300_peaks, hist_H3K4me3_peaks)
table(peaksOverlapsH3K4me3)
```
11130 of 24596 p300 peaks overlap with H3K4me3 marks.

Calculate proportion of overlap: true overlap of marks/total peaks:
```{r}
table(peaksOverlapsH3K4me3)[2]/(table(peaksOverlapsH3K4me3)[1]+table(peaksOverlapsH3K4me3)[2])
```
Proportion of overlap: ca. 39.3%


### H3K4me1

```{r}
peaksOverlapsH3K4me1 <- overlapsAny(tf_p300_peaks, hist_H3K4me1_peaks)
table(peaksOverlapsH3K4me1)
```
7919 of 24596 p300 peaks overlap with H3K4me1 marks.

Calculate proportion of overlap: true overlap of marks/total peaks:
```{r}
table(peaksOverlapsH3K4me1)[2]/(table(peaksOverlapsH3K4me1)[1]+table(peaksOverlapsH3K4me1)[2])
```
Proportion of overlap: ca. 27.9%

### H3K27ac

```{r}
peaksOverlapsH3K27ac <- overlapsAny(tf_p300_peaks,hist_H3K27ac_peaks )
table(peaksOverlapsH3K27ac)
```
16491 of 24596 p300 peaks overlap with H3K27ac marks.

Calculate proportion of overlap: true overlap of marks/total peaks:
```{r}
table(peaksOverlapsH3K27ac)[2]/(table(peaksOverlapsH3K27ac)[1]+table(peaksOverlapsH3K27ac)[2])
```
Proportion of overlap: ca. 58.1%

### H3K27me3

```{r}
peaksOverlapsH3K27me3 <- overlapsAny(tf_p300_peaks,hist_H3K27me3_peaks )
table(peaksOverlapsH3K27me3)
```
72 of 24596 p300 peaks overlap with H3K27me3 marks.

Calculate proportion of overlap: true overlap of marks/total peaks:
```{r}
table(peaksOverlapsH3K27me3)[2]/(table(peaksOverlapsH3K27me3)[1]+table(peaksOverlapsH3K27me3)[2])
```
Proportion of overlap: ca. 0.3%

